Revisiting Multi-Domain Machine Translation

### 2020-04-08 2020/coling2020/multidomain.tex
more todos, as a follow up of Talk with MinhQuang

In this paper, we try to answer these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet. Our main contribution is of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them. Our second contribution is experimental, and we perform a thorough reanalysis of most recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA. We show that\fyTodo{Spell out conclusions}.

@@

These various reasons for adopting MDMT are not mutually exclusive. Each however yields variable expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.

@@

For all models, we set the embeddings size and the size of hidden layers to~512. Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells. Some systems (see text below) additionally use an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024. The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
 %
Training uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$). In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers. % All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
\fyDone{Describe the block adaptation layer - voir slides}

\subsection{Multi-domain systems \label{ssec:systems}}
\fyDone{Remove Gated residual}
Our comparison of multidomain systems includes the following reimplementations of recent proposals from the literature:
\begin{itemize}
\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
\item the three proposals of \cite{Britz17mixing}: \system{TTM}, another tag-based approach where the domain tag is introduced as an extra word \textsl{on the target side}. Training is performed with reference tags and inference is performed with predicted tags, just like for any other target words; \system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations;
\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\fyTodo{URLs}); the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
\item An original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; residual connexions make it possible to by-pass these modules.
  % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
Note that contrarily to \cite{Bapna19simple} we do not start with a trained generic system, but learn the multi-domain from scratch. \fyDone{Check this.}
\end{itemize}

Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence, and slightly diverges from our notion of MDMT, which should be able to handle domain information in training and testing. This system has shown to meet many of our requirements, and would also have no problem scaling to a large number of domains.\fyTodo{TBC}

\section{Results and discussion \label{sec:results}}
\subsection{Performance of MDMT systems \label{ssec:rawperformance}}
In this section, we discuss the raw performance of MDMT systems trained and tested on 6~domains. Results are in Table~\ref{tab:performance}. As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2 BLEU points \jcDone{BP?} better, notably owing to the much better results for \domain{REL}). As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains. As all other systems are trained with an unbalanced data distribution, use the weighted average to perform global comparisons.

Fine-tuning each domain separately provides us with better baselines, significantly outperforming \system{generic} for all domains. A complete fine-tuning, which updates all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to lack of tuning the capacity of the adapter module. \fyDone{Add 2 averages ?}
A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word. All (except \system{DM} and \system{ADM}) improve over the generic baseline (for most domains), matching requirement [P1-LAB], but yield weaker results than fine tuning. As expected, the gap with fine-tuning remains large for an ``isolated'' domain such as \domain{rel}, the best contender being overall \system{MDL Res}, which approximately halves this gap as compared to \system{Mixed-Nat}.
\fyTodo{More comments when we have all the results}
The use of an decidated adaptation component in each of the transformer layer proves an even better strategy and closes the gap with fine-tuning for half of the domains: the use of a gating mechanism yielding a clear improvement over the board. The effect of the adaptation layer is especially significant for the smaller domains (\domain{bank}, \domain{IT} and \domain{rel}) for which unplugging this layer causes dramatic performance drops.

@@

    \system{DC-Tag}   & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5 &41.7 & 50.1\\
    \system{DC-Feat}  & 37.3   & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 &41.4 & 49.9\\
    \system{TTM}        & 37.7 & 54.8 & 49.7 & 33.4 & 43.8 & 80.0 &41.1 & 49.7\\
    \system{LDC-Tag} &&&&& &&&\\
    \system{DM}             & 36.5 & 52.4  & 47.7& 32.1 & 41.0 & 70.4 & 39.7 & 46.7 \\
    \system{ADM}           & 37.1 & 53.8  & 48.9 &32.8 & 42.3 & 78.5 & 40.7 & 48.9 \\
    \system{WDCMT}      & & & & & & & & \\
    \system{MDL Res}     & 37.9 & 56.0  & 51.2   & 33.5   &  44.4  & 88.3 & 42.1 & 51.9\\
    \system{MDL Res} (gen)           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3\\
     \hline
  \end{tabular}
  \caption{Translation performance of various MDMT systems. We report the BLEU score for each domain, as well as domain-weighted (wAVG) and unweighted (AVG) averages.}
  \label{tab:performance}
\end{table}

\subsection{Redefining domains \label{ssec:redomains}}
Table~\ref{fig:redomains} summaries the results of \fyTodo{XX} experiments where we artificially redefine the limits of domains. In the three \textsl{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario. This is the situation where we should see clear improvements even over finetuning, which only exploits part of the available relevant data. In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains. \fyTodo{Fix this}In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.

@@

\fyTodo{We do need significance testing}

@@

@@

Results are in table~\ref{tab:subdomains}.\fyTodo{Fill the table}

@@

% --------------
% \section*{Acknowledgements}

% The acknowledgements should go immediately before the references.  Do
% not number the acknowledgements section. Do not include this section
% when submitting your paper for review.

>>>>

In this paper, we try to answer these questions and to spell out in a more precise fashion the requirements that a MDMT system should meet. Our main contribution is of methodological nature and consists of lists of expected properties of MDMT system and associated measurements to evaluate them. \fyTodo{New problems - continuous learning, automatic domains}Our second contribution is experimental, and we perform a thorough reanalysis of most recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA. We show that\fyTodo{Spell out conclusions}.

@@

These various reasons for adopting MDMT are not mutually exclusive. Each however yields specific expectations with respect to the performance of this approach: if the motivation is primarily computational, then a drop in MT quality with respect to fine-tuning each domain might be acceptable; if it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some domains, if not all, over individually trained systems.

@@

For all models, we set the embeddings size and the size of hidden layers to~512. Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2048 cells. The multi-domain residual system (see description below) additionally use an adaptation block in each transformer layer, composed of 2-layer perceptron, with an inner RELU activation function operating on normalized entries of dimension 1024. The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
 %
Training uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4000$). In all cases, we use a batch size of~128 and a dropout rate of 0.1 for all layers. % All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.
\fyDone{Describe the block adaptation layer - voir slides}

\subsection{Multi-domain systems \label{ssec:systems}}
\fyDone{Remove Gated residual}
Our comparison of multi-domain systems includes the following reimplementations of recent proposals from the literature:
\begin{itemize}
\item a system using domain control as in \cite{Kobus17domaincontrol}: domain information is introduced either as an additional token for each source sentence (\system{DC-Tag}), or introduced in the form of a supplementary feature for each word (\system{DC-Feat}).
\item the three proposals of \cite{Britz17mixing}: \system{TTM}, another tag-based approach where the domain tag is introduced as an extra word \textsl{on the target side}. Training is performed with reference tags and inference is performed with predicted tags, just like for any other target words; \system{DM} is a multi-task learner where a domain classifier is trained on top the MT encoder, so as to make it aware of domain differences; \system{ADM} is the adversarial version of \system{DM}, pushing the encoder towards learning domain-independent source representations;
\item a system using lexicalized domain representations \cite{Pham19generic}: word embeddings are composed of a generic and a domain specific part (\system{LDC-Feat});
\item the multi-domain model of \cite{Zeng18multidomain} (\system{WDCMT}): in this approach two representations, a domain-agnostic and a domain-specialized representation of the input are simultaneously processed; supervised classification models and adversarial training are used to compute these representations.\footnote{For this system, we use the available RNN-based system from the authors (\fyTodo{URLs}) which does not directly compare to the other, Transformer-based, systems; the improved version of \cite{Su19exploring} seems to produce comparable, albeit slighly improved version.}
\item An original, multi-domain version of the approach of \newcite{Bapna19simple}, denoted \system{MDL Res}, where we add a dedicated adaptation module for each domain, which is only activated for the appropriate set of training data; within each layer, residual connexions make it possible to by-pass this module.
  % In a variant (\system{MDL Gated}), we use a gating mechanism to merge computations using the the adapter modules and with those that don not.
Note that contrarily to \cite{Bapna19simple} we do not start with a trained generic system, but learn the multi-domain from scratch. \fyDone{Check this.}
\end{itemize}

Note that we did not include here the system of \cite{Farajian17multidomain} which performs on-the-fly tuning for each test sentence, and slightly diverges from our notion of MDMT, which should be able to handle domain information in training and testing. This system has shown to meet many of our requirements and would also have no problem scaling to a large number of domains.\fyTodo{TBC}

\section{Results and discussion \label{sec:results}}
\subsection{Performance of MDMT systems \label{ssec:rawperformance}}
In this section, we discuss the raw performance of MDMT systems trained and tested on $6$~domains. Results are in Table~\ref{tab:performance}. As expected, balancing data in the generic setting makes a great difference (the unweighted average is 2~BLEU points \jcDone{BP?} better, notably owing to the much better results for \domain{REL}). As explained above, this setting should be the baseline when the test distribution is assumed to be balanced across domains. As all other systems are trained with an unbalanced data distribution, use the weighted average to perform global comparisons.

Fine-tuning each domain separately provides us with better baselines, significantly outperforming \system{generic} for all domains. A complete fine-tuning, which updates all parameters, proves more effective than only adapting a subpart of the network - a result that could be due to lack of tuning the capacity of the adapter module. \fyDone{Add 2 averages ?}\fyTodo{Significance testing wrt Mix Generic, Full Fine-tuned, for each domain}
A  first group of MDMTs use an extra domain feature, at the sentence or at the word level: among these, having the domain tag on the source is better than on the target, and very comparable to repeating the tag for every source word. All (except \system{DM} and \system{ADM}) improve over the generic baseline (for most domains), matching requirement [P1-LAB], but yield weaker results than fine tuning. As expected, the gap with fine-tuning remains large for an ``isolated'' domain such as \domain{rel}, the best contender being overall \system{MDL Res}, which approximately halves this gap as compared to \system{Mixed-Nat}.
\fyTodo{More comments when we have all the results}
The use of an decidated adaptation component in each of the transformer layer proves an even better strategy and closes the gap with fine-tuning for half of the domains: the use of a gating mechanism yielding a clear improvement over the board. The effect of the adaptation layer is especially significant for the smaller domains (\domain{bank}, \domain{IT} and \domain{rel}) for which unplugging this layer causes dramatic performance drops.

@@

    \system{DC-Tag}   & 38.1 & 55.3 & 49.9   & 33.2 & 43.5 & 80.5 &41.7 & 50.1\\
    \system{DC-Feat}  & 37.3   & 54.9 & 49.5   & 32.9 & 43.6 & 79.9 &41.4 & 49.9\\
    \system{TTM}        & 37.7 & 54.8 & 49.7 & 33.4 & 43.8 & 80.0 &41.1 & 49.7\\
    \system{LDC-Tag} &&&&& &&&\\
    \system{DM}             & 36.5 & 52.4  & 47.7& 32.1 & 41.0 & 70.4 & 39.7 & 46.7 \\
    \system{ADM}           & 37.1 & 53.8  & 48.9 &32.8 & 42.3 & 78.5 & 40.7 & 48.9 \\
    \system{WDCMT}      & & & & & & & & \\
    \system{MDL Res}     & 37.9 & 56.0  & 51.2   & 33.5   &  44.4  & 88.3 & 42.1 & 51.9\\
    \system{MDL Res} (gen)           & 37.7 & 51.0 & 34.0 & 30.4 & 34.2 & 15.2 & 36.4 & 33.7\\
%    \system{MDL Gated} & 37.7 & 56.5 & 53.2 & 34.1 & 44.6 & 90.7 & 42.3 & 53.3\\
     \hline
  \end{tabular}
  \caption{Translation performance of various MDMT systems. We report the BLEU score for each domain, as well as domain-weighted (wAVG) and unweighted (AVG) averages.}
  \label{tab:performance}
\end{table}
\fyTodo{Fill the table with all results, including a comparison with *RNNs* for wcmd}
\ftTodo{A new column for the News domain with predicted domains (except for the generic case).}

\subsection{Redefining domains \label{ssec:redomains}}
Table~\ref{fig:redomains} summaries the results of \fyTodo{XX} experiments where we artificially redefine the limits of domains. In the three \textsl{split} experiments, we randomly split a domain corpus in two part: a good MDMT system should detect that these two ``domains'' should be mutually beneficial, and should be hardly affected by this change wrt.\ the baseline scenario. This is the situation where we should see clear improvements even over finetuning, which only exploits part of the available relevant data. In the \textsl{merge} experiment, we do the reverse and merge two corpora in training, in order to assess the robustness with respect to heterogenous domains. \fyTodo{Prepare table for the wrong domain analysis (see also last col. of the first table)}In the last column, we finally report the averaged (across domains) drop in performance when the domain information is erroneous.

@@

\fyTodo{We do need significance testing (a) each system against itself (non-split); (b) each system agains FT in split  condition}

@@

\fyTodo{Instert table, significancy testing against full learning condition, for each domain}

@@

Results are in table~\ref{tab:subdomains}.\fyTodo{Fill the tables with automatic and existing domains - we will have significancy results for the existing domains, for the other may be for larger classes we will have enough test instances?}

@@

% --------------
% \section*{Acknowledgements}
% This work was granted access to the HPC resources of [TGCC/CINES/IDRIS] under the allocation 20XX- [numéro de dossier] made by GENCI (Grand Equipement National de Calcul Intensif)
% The acknowledgements should go immediately before the references.  Do
% not number the acknowledgements section. Do not include this section
% when submitting your paper for review.
